---
title: "In Class Exercise 5"
author: "Arjun Singh"
date: 2024-09-23
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
format:
  html:
    css: styles.css  
---

# 5 Introduction

This exercise will help reinforce our learning from Hands on exercise 5. Additionally, we explore the sfdep package and its uses.

# 5.1 Objective

In spatial policy, a key objective for local governments and urban planners is to promote balanced development across the province. In this study, our goal is to apply appropriate spatial statistical methods to assess whether development is evenly distributed geographically in Hunan Province, People's Republic of China, using GDP per capita as a development indicator.

If we find that development is not evenly distributed, our next step will be to investigate whether there are signs of spatial clustering. Should clustering be present, we will then identify the locations of these clusters. Through this exercise, we aim to uncover the spatial patterns of development in the province.

# 5.2 Data and Packages

For this exercise, we have the following two datasets:

-   Hunan province administrative boundary layer at county level. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

The packages used are as follows:

-   **`sf`**: Provides simple features support for handling and analyzing spatial vector data in R.
-   **`spdep`**: A package for spatial dependence and spatial regression analysis, particularly for handling spatial weights.
-   **`tmap`**: A flexible visualization package for thematic maps, supporting both static and interactive mapping in R.
-   **`tidyverse`**: A collection of R packages designed for data science, emphasizing data manipulation, visualization, and functional programming.
-   `sfdep` is designed for spatial analysis, providing tools to work with spatial dependencies and spatial econometrics, particularly in conjunction with simple features (sf) objects. It helps analyze relationships and interactions between spatially distributed data.

We import these packages into our environment using the following code chunk.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, sfdep)
set.seed(1234)
```

# 5.3 Importing the Data

::: panel-tabset
## Geospatial Data

We will use the st_read() function of the sf package to import the Hunan Shapefile into our environment.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

## Aspatial Data

We will use the read_csv() function of the readr package to import the Hunan_2012 data file.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")

```
:::

## 5.3.1 Performing relational join

We will update the attribute table of `Hunan`'s SpatialPolygonsDataFrame with the attribute fields of the `hunan2012` data-frame. We can do this by using the left_join() function of the dplyr package.

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### 5.3.1.2 Plotting a Choropleth map

```{r}
tmap_mode('plot')
tm_shape(hunan_GDPPC)+
  tm_fill('GDPPC', style='quantile', palette = 'Blues', title= 'GDPPC')
```

# 5.4 Global and Local Measures of Spatial Autocorrelation: sfdep methods

## 5.4.1 Deriving Queens Contiguity Weights: sfdep methods

```{r}
wm_q=hunan_GDPPC%>%
  mutate(nb=st_contiguity(geometry),
         wt=st_weights(nb,
                       style='W'),
         .before=1)
```

-   nb: A neighbor list object as created by st_neighbor.

-   style: Default 'W' for row standardized weights. This value can also be 'B', 'C', 'U', 'minimax', and 'S'

-   allow_zero: if TRUE, assigns zero as a lagged value to zone without neighbors.

## 5.4.2 Computing Global Moran's I

We use the global_moran() function to compute this.

```{r}
moranI= global_moran(wm_q$GDPPC,
                     wm_q$nb,
                     wm_q$wt)
```

The below code chunk, using global_moran_test() helps us conduct the test easier.

```{r}
global_moran_test(wm_q$GDPPC,
                  wm_q$nb,
                  wm_q$wt)
```

Since p-value is less than 0.05, we infer that there is indeed a sign of positive autocorrelation.

## 5.4.3 Global Moran's Permutation Test

In practice, we generally use Monte Carlo simulation to conduct tests. This is where the global_moran_perm() function comes into play. The `nsim` argument is key.

```{r}
global_moran_perm(wm_q$GDPPC,
                  wm_q$nb,
                  wm_q$wt,
                  nsim = 99)
```

## 5.4.4 Computing Local Moran's I

In this section, we use the local_moran() function of the sfdep package.

```{r}
lisa=wm_q%>%
  mutate(local_moran=local_moran(
    GDPPC, nb, wt, nsim=99),
    .before=1)%>%
  unnest(local_moran)
```

unnest() helps us combine the data into the intended data-frame.

## 5.4.5 Visualizing p-value of local Moran's I

```{r}
tmap_mode('plot')
tm_shape(lisa)+
  tm_fill('p_ii_sim')+
  tm_borders(alpha=0.5)+
  tm_layout(main.title = 'p-value of Local Morans I',
            main.title.size = 2)
```

### 5.4.5.1 Visualizing Local Moran's I

```{r}
tm_shape(lisa)+
  tm_fill('ii')+
  tm_borders(alpha=0.5)+
  tm_view(set.zoom.limits = c(6,8))+
  tm_layout(main.title='Local Morans I of GDPPC',
            main.title.size = 2)
```

### 5.4.5.2 Visualizing both

```{r}
map1= tm_shape(lisa)+
  tm_fill('p_ii_sim')+
  tm_borders(alpha=0.5)+
  tm_layout(main.title = 'p-value of Local Morans I',
            main.title.size = 2)

map2= tm_shape(lisa)+
  tm_fill('ii')+
  tm_borders(alpha=0.5)+
  tm_view(set.zoom.limits = c(6,8))+
  tm_layout(main.title='Local Morans I of GDPPC',
            main.title.size = 2)

tmap_arrange(map1,map2, ncol=2)
```

## 5.4.6 Visualizing LISA map

LISA map is a categorical map showing outliers and clusters. There are two types of outliers- high-low and low-high outliers. Likewise, there are two types of clusters namely High-High and Low-Low.

In fact, LISA map is an interpreted map by combining local Moran's I of geographical areas and their respective p values.

```{r}
lisa_sig=lisa%>%
  filter(p_ii<0.05)
tmap_mode('plot')
tm_shape(lisa)+
  tm_polygons()+
  tm_borders(alpha=0.5)+
  tm_shape(lisa_sig)+
  tm_fill('mean')+
  tm_borders(alpha=0.4)
```

We see two outliers in the cluster on the North-East region of Hunan which contains mostly high-high regions.

# 5.5 Computing Local Gi statistics

```{r}
wm_idw=hunan_GDPPC%>%
  mutate(nb=st_contiguity(geometry),
         wts=st_inverse_distance(nb, geometry, scale=1, alpha=1),
         .before=1)
```

```{r}
HCSA=wm_idw%>%
  mutate(local_Gi=local_gstar_perm(
    GDPPC, nb, wt, nsim=99),
    .before=1)%>%
  unnest(local_Gi)
HCSA
```

```{r}
tm_shape(HCSA)+
  tm_fill('p_sim')+
  tm_borders(alpha=0.5)
```

We can now visualize the hot and cold spots.

```{r}
HCSA_sig=HCSA%>%
  filter(p_sim<0.05)
tm_shape(HCSA)+
  tm_polygons()+
  tm_borders(alpha = 0.5)+
  tm_shape(HCSA_sig)+
  tm_fill('cluster')
  tm_borders(alpha=0.4)
```
