---
title: "In Class Exercise 4"
author: "Arjun Singh"
date: 2024-09-16
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
format:
  html:
    css: styles.css
---

# Introduction

In this exercise, we focus on reinforcing our learning from Hands-on Exercise 4.

Additionally, we also implement the `GWModel` package in this exercise.

The **`GWModel`** package in R is designed for performing geographically weighted (GW) models, which account for spatial heterogeneity by allowing local rather than global parameter estimates. This is particularly useful when the relationship between your dependent and independent variables might vary across space, such as in spatial regression analysis.

# Data and Packages

For this exercise, we will be focusing on Hunan. The following data sets will be used:

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

The following packages are used:

-   **`sf`**: Provides simple features support for handling and analyzing spatial vector data in R.

-   **`spdep`**: A package for spatial dependence and spatial regression analysis, particularly for handling spatial weights.

-   **`tmap`**: A flexible visualization package for thematic maps, supporting both static and interactive mapping in R.

-   **`tidyverse`**: A collection of R packages designed for data science, emphasizing data manipulation, visualization, and functional programming.

-   **`knitr`**: A dynamic report generation tool in R, allowing for the integration of code, results, and narrative in reproducible documents.

-   `GWmodel` has been discussed above, above this section.

-   `ggstatsplot` is a power visualization package. It enhances `ggplot2` by including the results of hypothesis tests, effect sizes and confidence intervals on the plots.

    We import these packages into our environment.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel, ggstatsplot)
set.seed(123)
```

# 4.2 Importing the data

We start of by importing the Hunan shapefile into R using the st_read() function of the sf package.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

We see that it is a simple feature data-frame with 88 features and 7 fields, projected in the WGS84 coordinate system.

We now import the csv file using the read_csv() function of the readr package.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

This is a simple R data-frame.

## 4.2.1 Performing relational joins

We will proceed to join this to the simple feature data-frame, `hunan`, obtained above using a relational join technique.

The left_join() function of the dplyr package is used as shown in the code chunk below.

```{r}
hunan <- left_join(hunan,hunan2012)%>%
  dplyr::select(1:3, 7, 15, 16, 31, 32) # This facilitates efficiency... dropping unnecessary columns is important.
```

# 4.9 In-class Exercise: GWmodel Package

We specifically focus on the **gwss** function of the GWmodel package.

::: insights-box
The `gwss()` function in the **`GWModels`** R package stands for **Geographically Weighted Summary Statistics**. It allows you to calculate local summary statistics for a set of variables, such as mean, standard deviation, skewness, and more, at different spatial locations. This is useful for understanding spatial variability in your data.
:::

In order to use a GWmodel, we must use the as_Spatial() function of the sf package in order to convert our simple feature data-frame into a spatial object.

```{r}
hunan_sp=hunan%>%
  as_Spatial()
```

## 4.9.1 Geographically Weighted Summary Statistics with adaptive bandwidth

### 4.9.1.1 Determine Adaptive Bandwidth

```{r}
bw_CV=bw.gwr(GDPPC ~ 1, #arbitrary number
             data= hunan_sp,
             approach = 'CV',
             adaptive = TRUE,
             kernel = 'bisquare',
             longlat = T)
```

```{r}
bw_CV
```

::: note-box
Note: We can also use the AIC method.

```{r}
bw_AIC=bw.gwr(GDPPC ~ 1, #arbitrary number
             data= hunan_sp,
             approach = 'AIC',
             adaptive = TRUE,
             kernel = 'bisquare',
             longlat = T) #Converts everything into Kilometers.
```

```{r}
bw_AIC
```
:::

### 4.9.1.2 Determining Fixed Bandwidth

In this case, the code generally remains the same except one key difference. We set the adaptive argument to FALSE.

::: panel-tabset
## CV

```{r}
bw_fixed_CV=bw.gwr(GDPPC ~ 1, #arbitrary number
             data= hunan_sp,
             approach = 'CV',
             adaptive = FALSE,
             kernel = 'bisquare',
             longlat = T)
```

## AIC

```{r}
bw_fixed_AIC=bw.gwr(GDPPC ~ 1, #arbitrary number
             data= hunan_sp,
             approach = 'AIC',
             adaptive = FALSE,
             kernel = 'bisquare',
             longlat = T)
```
:::

We notice a very significant difference between fixed and adaptive bandwidth approaches.

## 4.9.2 Computing Geographically Weighted Summary Statistics

We will now implement the gwss() function to compute the summary statistics.

```{r}
gwstat=gwss( data = hunan_sp,
             vars='GDPPC',
             bw=bw_AIC,
             kernel = "bisquare",
             adaptive=TRUE,
             longlat = T)
gwstat
```

::: note-box
Note: the Kernel and Adaptive must be consistent for computation of bandwidth as well as the Summary Statistics.
:::

## 4.9.3 Preparing the output data

We implement the code chunk below to extract the SDF data table from the gwss object output fro gwss(). It is converted into a data-frame by applying the as.data.frame() function.

```{r}
gwstat_df=as.data.frame(gwstat$SDF)

```

After this, we can implement cbind() to append the newly derived data-frame onto the `hunan` sf data frame.

```{r}
hunan_gstat=cbind(hunan, gwstat_df)
```

## 4.9.4 Visualization

We will now plot the above obtained data-frame in order to facilitate visualization.

```{r}
tm_shape(hunan_gstat)+
  tm_fill('GDPPC_LM',
          n=5,
          style='quantile')+
  tm_borders(alpha = 0.5)+
  tm_layout(main.title = "Distribution of geographically weighted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            legend.text.size = 1.2,
            legend.height = 1.5,
            legend.width = 1.5,
            frame = TRUE)
```

## 4.9.5 Geographically weighted correlation with adaptive bandwidth

### 4.9.5.1 Computing Correlation

```{r}
bw=bw.gwr(GDPPC ~ GIO, #arbitrary number
             data= hunan_sp,
             approach = 'AIC',
             adaptive = TRUE,
             kernel = 'bisquare',
             longlat = T)
```

```{r}
gwstats=gwss(hunan_sp,
             vars=c('GDPPC', 'GIO'),
             bw=bw,
             kernel='bisquare',
             adaptive = TRUE,
             longlat=T)
             
```

We will now plot the scatterplot with the best fit line as well as required summary statistics to further our understanding of the data.

```{r}

```

::: panel-tabset
## Local Correlation Coefficient

```{r}

```

## Local Spearman Coefficient

```{r}

```
:::
