---
title: "In Class Exercise 10"
author: "Arjun Singh"
date: 2024-11-4
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
format:
  html:
    css: styles.css 
---

# Introduction

In this in-class exercise, we focus on the following:

-   Preparing data downloaded from REALIS portal for geocoding,
-   Geocoding by using SLA OneMap API,
-   Converting the geocoded transaction data into sf point feature data.frame, and
-   Wrangling the sf point features to avoid overlapping point features.

## Data and Packages

::: {style="font-size: 1.50em"}
```{r}
pacman::p_load(tidyverse, sf, tmap, httr, performance)
```
:::

## Importing data

The code chunk below imports multiple csv files in a specified folder and append them into a single tibble data frame.

::: {style="font-size: 1.50em"}
```{r}
folder_path <- "data/aspatial"
file_list <- list.files(path = folder_path, 
                        pattern = "^Residential.*\\.csv$", 
                        full.names = TRUE)

realis_data <- file_list %>%
  map_dfr(read_csv)
```
:::

## Wrangling data

::: panel-tabset
### The task

Write a code chunk to perform the followings: - converting values in *Sale Date* field from character to numerical date format, and - extracting resale and condominium transaction records.

### The code

::: {style="font-size: 1.50em"}
```{r}
condo_resale <- realis_data %>%
  mutate(`Sale Date` = dmy(`Sale Date`)) %>%
  filter(`Type of Sale` == "Resale" &
           `Property Type` == "Condominium")
```
:::
:::

## Geocoding

::: panel-tabset
### Preparing data

::: {style="font-size: 1.50em"}
```{r}
postcode <- unique(condo_resale$`Postal Code`)
```
:::

### Geocoding

::: {style="font-size: 1.50em"}
```{r}
url <- "https://onemap.gov.sg/api/common/elastic/search"
found <- data.frame()
not_found <- data.frame()

for (postcode in postcode){
  query <- list('searchVal'=postcode, 'returnGeom'='Y', 
                'getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url, query=query)
  if ((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else {not_found = data.frame(postcode)
  }
}
```
:::

```{r}
#| echo: false
write_rds(found, "data/rds/found.rds")
```
:::

## Tidying field names

::: {style="font-size: 1.50em"}
```{r}
found <- found %>%
  select(c(6:8)) %>%
  rename(POSTAL = `results.POSTAL`,
         XCOORD = `results.X`,
         YCOORD = `results.Y`)
```
:::

## Converting to Point Feature Data Frame

::: panel-tabset
### The tasks

-   Write a code chunk to join *condo_resale* and *found*. Name the output *condo_resale_geocoded*.

-   Write a code chunk to convert *condo_resale_geocoded* from tibble data frame to sf point feature data frame.

### Joining tables

::: {style="font-size: 1.50em"}
```{r}
condo_resale_geocoded = left_join(
  condo_resale, found, 
  by = c('Postal Code' = 'POSTAL'))
```
:::

### Convering to sf

::: {style="font-size: 1.50em"}
```{r}
condo_resale_sf <- st_as_sf(condo_resale_geocoded, 
                            coords = c("XCOORD",
                                       "YCOORD"),
                            crs=3414)
```
:::
:::

## Cleaning Spatial Data

::: panel-tabset
### Checking for overlapping point features

The code chunk below is used to check if there are overlapping point features.

::: {style="font-size: 1.50em"}
```{r}
overlapping_points <- condo_resale_sf %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```
:::

### Spatial jittering

In the code code chunk below, `st_jitter()` of sf package is used to move the point features by 5m to avoid overlapping point features.

::: {style="font-size: 1.50em"}
```{r}
condo_resale_sf <- condo_resale_sf %>%
  st_jitter(amount = 2)
```
:::
:::
