---
title: "In Class Exercise 2"
author: "Arjun Singh"
date: 2024-09-02
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# Introduction

In this exercise, we reinforce our learning and deal with any problems we faced from Hands-On Exercise 2 by further application of the skills gained from it.

# Data and Packages

For this exercise, we will use the following R packages:

-   **sf**, a relatively new R package specially designed to import, manage and process vector-based geospatial data in R.

-   **spatstat**, which has a wide range of useful functions for point pattern analysis. In this hands-on exercise, it will be used to perform 1st- and 2nd-order spatial point patterns analysis and derive kernel density estimation (KDE) layer.

-   **raster**, which reads, writes, manipulates, analyses and model of gridded spatial data (i.e. raster). In this hands-on exercise, it will be used to convert image output generate by spatstat into raster format.

-   **tidyverse** simplifies spatial analysis by offering a consistent and efficient framework that facilitates working with spatial data.

-   **tmap** which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using leaflet API.

We will load in the packages above by implementing the code chunk below.

<details>

<summary>Click to show/hide code</summary>

```{r}
pacman::p_load(sf, spatstat, raster, tmap, tidyverse)
set.seed(1234)
```

</details>

To install the **maptools** package, which is now retired, we will use the code chunk below.

```{r}
#| eval: false
install.packages("maptools", repos="https://packagemanager.posit.co/cran/2023-10-13") #Posit is a great source for downloading retired libraries, and even facilitates the development of Shiny apps.
```

We will now import the *GEOJSON* using the **st_read()** function of the **sf** package.

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

We notice that its projected in the **WGS84** system so we transformed the **EPSG** code to **3414**, to be consistent with the **SVY21** projected coordinate system of Singapore.

We will now import the coastal outline data using the **st_read()** function again.

```{r}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")
```

This is projected correctly in the SVY21 system, however we will verify the EPSG code using the **st_crs()** function of the sf package.

```{r}
st_crs(sg_sf)
```

We notice that the EPSG code isn't correct, so we transform it to the correct value of 3414 using the **st_transform()** function.

```{r}
st_transform(sg_sf, 3414)
```

Now, we import the Master Planning Subzone data using the **st_read()** function

```{r}
mpsz_sf <- st_read(dsn = "data", 
                   layer = "MP14_SUBZONE_WEB_PL")
```

We must now verify the EPSG code for `mpsz_sf` before proceeding to facilitate analysis.

```{r}
st_crs(mpsz_sf)
```

Since it has the wrong EPSG code, we will transform it to the correct value of 3414.

```{r}
st_transform(mpsz_sf, 3414)
```

Now, we can proceed with the analysis as all the data-frames are projected consistently.

# Creating a coastal outline

Using the **st_union()** function, we can derive the coastal outline of the **sf tibble data-frame**.

```{r}
sg_sf= mpsz_sf %>% st_union
```

We will then have the figure below, which will help improve the quality of our analysis.

```{r}
plot(sg_sf)
```

# Kernel Density Estimation

We now implement two different methods to convert KDE output into grid object.

We first convert the data into the right format for analysis.

```{r}
childcare_ppp=as.ppp(childcare_sf)
childcareSG_ppp.km <- rescale.ppp(childcare_ppp, 1000, "km")
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
```

::: panel-tabset
## maptools method

```{r}
gridded_kde_childcareSG_ad=maptools::as.SpatialGridDataFrame.im(kde_childcareSG_adaptive)
plot(gridded_kde_childcareSG_ad)

```

## spatstat.geom method

```{r}
gridded_kde_childcareSG_ad=as(kde_childcareSG_adaptive, "SpatialGridDataFrame")
plot(gridded_kde_childcareSG_ad)
```
:::

As we can see, both have very similar/same results however we prefer to use spatstat.geom as maptools has been retired..

# Importing the Traffic Accident Data

We will import the data file into our environment.

```{r}
rdacc_sf=read_csv("data/thai_road_accident_2019_2022.csv")%>%
  filter(!is.na(longitude) & longitude != "",
         !is.na(latitude) & latitude !="")%>%
  st_as_sf(coords=c(
    "longitude", "latitude"), 
    crs=4326)%>%
  st_transform(crs=32647)
```

::: panel-tabset
## Visualizing the accident data.
:::
